A Container-Based Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms

### 2020-10-14 draft.tex
Fixed grammar issues
These fixes were highlithed by reviewer #4 it closes #19

Researchers 
and algorithm designers willing to  adapt to the cloud their
multi-population solutions face the additional challenges:
\begin{itemize}
    \item They need to change their current solutions to a reactive architecture in which processes
    communicate with each other by exchanging messages asynchronously and react
    to a continuous stream of data. One of the options is to consider
    populations as messages to be modified asynchronously by functions.

@@

%%%%%% -------------------- This too. Should go up, to intro, or down,
%%%%%% to method - JJ
Some parameters, specially the population size, are
kept fixed in order to control more easily the execution of the algorithm. For
instance, by having the size of populations fixed, it is easier to control
the number of evaluations and the communication costs, when the algorithm is in
operation. % And to keep them more or less synchronized, which can be
           % interesting, too... But this should go to the
           % introduction also - JJ

@@

The most elemental data structure used throughout the system is the population; and at a higher
level, we can view the system as a continuous stream of populations
flowing from one process to the other. By ``continuous stream,'' we mean that ideally,
everything must happen asynchronously without components needing to wait for
others. In this kind of system, this streaming functionality is normally
implemented by using a message queue system. For instance, if one process needs
to communicate a population to another, it must pack the population into
a message, and then send the message 
to the queuing system without waiting for a response; this
means that after the process pushes the message, it continues its execution
without waiting for a result. On the other end, the recipient subscribes to the
message queue by defining an event handler method that is going to be triggered when a
message is pushed to it. The two components at each end of the queue are
Producers that push messages to the queue and Consumers that pull them. This
pattern is an essential component of a reactive architecture and makes
it highly scalable; one of the reasons for this is the use of functions
that have no secondary effects or access global storage, that is,
stateless functions. These functions do not need to read, keep, or modify data outside of the method. If
Consumers are implemented using stateless functions, and thus, there is no difference
between having one or many copies of the same function pulling work from the queue all at the
same time, because no side effects, including unwanted resource
locking problems,  are going to result from this concurrency.
% This is important and we should probably mention that in the
% introduction - JJ 
Based on these general principles we define next the proposed model.

@@

EvoSwarm describes the deployment of loosely coupled processes in a consistent
and portable way. The term cloud-native describes these container-based
environments because we can deploy them on an infrastructure that abstracts the
underlying compute, storage, and networking primitives. This infrastructure does
not need to be a cloud provider; it could be a developer's laptop, a single
server, or a production environment. In this section, we give specific details
about how we deployed EvoSwarm in a cloud-setting for running the experiments
within this paper. Running the cloud-native application in the cloud has the obvious
benefit of a multi-instance deployment, where we can run containers in several
instances executing at the same time, not only in a single machine, not only 
in a single machine as it is the case when using docker-compose.

@@

Figure \ref{fig:aws-configuration} shows the basic EC2 configuration used for the deployment
of experiments. We selected three types of EC2 instances according to our needs; the  
{\texttt setup}, {\texttt controller}, and {\texttt redis} containers do not demand many resources
and we do not need to scale them as the amount of parallel workers are increased. They run in a
single {\texttt C5  High-CPU large} instance. With the available 2048 CPU shares divided as follows:
{\texttt setup} (128),  {\texttt controller} (896),{\texttt redis} (1024). With this assignment 
the {\texttt redis} container gets a dedicated vCPU, needed because it is the most demanding 
task of the three. We assigned 2048 units to each {\texttt worker} container, and using a more 
capable {\texttt C5 High-CPU XLarge} instance,  with 
four vCPUs and 4096 CPU shares, with this configuration each worker gets a pair of vCPUs.  
We scaled the number if instances as needed, using one instance for every two {\texttt worker} containers.

>>>>

Researchers  and algorithm designers willing to adapt to the cloud; however, their
multi-population solutions face additional challenges:
\begin{itemize}
    \item They need to change their current solutions to a reactive architecture in which processes
    communicate with each other by exchanging messages asynchronously and react
    to a continuous stream of data. One of the options is to consider
    populations as messages to be modified asynchronously by functions.

@@

%%%%%% -------------------- This too. Should go up, to intro, or down,
%%%%%% to method - JJ
Some parameters, especially the population size, are
kept fixed in order to control the execution of the algorithm more easily. For
instance, by having the size of populations fixed, it is easier to control
the number of evaluations and the communication costs, when the algorithm is in
operation. % And to keep them more or less synchronized, which can be
           % interesting, too... But this should go to the
           % introduction also - JJ

@@

The most elemental data structure used throughout the system is the population; and at a higher
level, we can view the system as a continuous stream of populations
flowing from one process to the other. By ``continuous stream,'' we mean that ideally,
everything must happen asynchronously without components needing to wait for
others. In this kind of system, this streaming functionality is normally
implemented by using a message queue system. For instance, if one process needs
to communicate a population to another, it must pack the population into
a message, and then send the message 
to the queuing system without waiting for a response; this
means that after the process pushes the message, it continues its execution
without waiting for a result. On the other end, the recipient subscribes to the
message queue by defining an event handler method that is going to be triggered when a
message is pushed to it. The two components at each end of the queue are
Producers that push messages to the queue and Consumers that pull them. This
pattern is an essential component of a reactive architecture and makes
it highly scalable; one of the reasons for this is the use of functions
that have no secondary effects or access global storage, that is,
stateless functions. These functions do not need to read, keep, or modify data outside of the method. If
Consumers are implemented using stateless functions, and thus, there is no difference
between having one or many copies of the same function pulling work from the queue all at the
same time, because no side effects, including unwanted resource
locking problems,  are going to result from this concurrency.
% This is important and we should probably mention that in the
% introduction - JJ 
Based on these general principles, we next define the proposed model.

@@

EvoSwarm describes the deployment of loosely coupled processes in a consistent
and portable way. The term cloud-native describes these container-based
environments because we can deploy them on an infrastructure that abstracts the
underlying compute, storage, and networking primitives. This infrastructure does
not need to be a cloud provider; it could be a developer's laptop, a single
server, or a production environment. In this section, we give specific details
about how we deployed EvoSwarm in a cloud-setting for running the experiments
within this paper. Running the cloud-native application in the cloud has the obvious
benefit of a multi-instance deployment, where we can run containers in several
instances executing at the same time, not only in a single machine as it is the case when using docker-compose.

@@

Figure \ref{fig:aws-configuration} shows the basic EC2 configuration used for the deployment
of experiments. We selected three types of EC2 instances according to our needs; the  
{\texttt setup}, {\texttt controller}, and {\texttt redis} containers do not demand many resources
and we do not need to scale them as the amount of parallel workers are increased. They run in a
single {\texttt C5  High-CPU large} instance. With the available 2048 CPU shares divided as follows:
{\texttt setup} (128),  {\texttt controller} (896), {\texttt redis} (1024). With this assignment 
the {\texttt redis} container gets a dedicated vCPU, needed because it is the most demanding 
task of the three. We assigned 2048 units to each {\texttt worker} container, and using a more 
capable {\texttt C5 High-CPU XLarge} instance,  with 
four vCPUs and 4096 CPU shares, with this configuration each worker gets a pair of vCPUs.  
We scaled the number of instances as needed, using one instance for every two {\texttt worker} containers.
